---
title: "index.rmd"
output: html_document
date: '2022-05-16'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
```

#### Preliminary Note

Please note that you will need to run all script files prior to running this file, such that you have all the required objects saved to your environment.

## Introduction

This report studies the visitation of the United States National Parks Service. We break down based on state and region to determine which state makes the best use of their national park(s). We include data about the size of each park to compare if larger parks get more visitors, and also to see what the ratio of park size to visitors is (ie which park earns the most "bang for its buck). Parks that are smaller and have higher visitation are more efficient and garner more business for the NPS. These parks however are not to be exalted in comparison to those with a very low visitor to area ratio (meaning there are many fewer visitors per acre of land). We claim that the parks with the lowest ratio of visitors to area are the most successful at maintaining and untouched landscape. These parks take large areas and keep them the most clear of human presence.

## Summary Information

```{r summary_calc}
View(summary_info)
```

### Summary Info Component Description

- num_parks: represents the number of parks present in the data
- largest_park: represents the park which encompasses the largest land area
- smallest_park: represents the park which encompasses the least land area
- most_visited_state: shows the state that had the most visitors to any of its national parks between 2010 and 2016
- least_visited_state: shows the state that had the fewest visitors in the timeframe (Note: American Samoa is not a state, rather a territory. It does contain a national park though, so it is included)
- most_visited_region: the region with the most visitors over the timeframe
- least_visited_region: the region with the least visitors over the timeframe (Note: Alaska is its own region)
- most_visited_park_for_year: the park which had the most visitors accross the whole dataframe, and the year it took place
- least_visited_park_for_year: the park with the fewest visitors accross the data frame, and the year it occured. Note: there was a tie for this, so both instances are included.
- most_visited_per_acre: shows the park with the highest visitors per acre, and the ratio itself
- least_visited_per_acre: shows the park with the least visitors per acre, and the ratio itself

## Table

```{r display_table}
View(state_np_area)
View(vis_by_state)
View(vis_to_acre)
```

### Table Description

##### State NP Area

The state_np_area table shows each state that has a national park, and lists the total area that is in that state. The purpose here is to show which states have allocated the most land towards National Parks

##### Visitation by State

The vis_by_state table shows each state's total visitation from 2010-2016. This is not necessarily correlated with the state area table, but does show which states have the msot visited National Park sites.

##### Visitation per Acre

The vis_to_acre table shows each national park's millions of visitors to acres ratio. For example, a ration of 12 would indicate that over the 6 year time period, there were 12 million visitors per 1 acre of that national park.

## Charts

### Chart 1

```{r chart1}
# insert chart1
```

### Chart 2

```{r chart2}
# insert chart 2
```

### Chart 3

```{r chart3}
chart_3
```

This chart is  meant to show the visitation to the entire NP system across the time period to determine if more people are visiting more recently. Overall, the number seems to be increasing, with a big outlier in 2012, but still a notable positive trend.
